---
title: "ST 558 Project 1"
author: "John Hinic"
date: '2022-06-26'
output:
  pdf_document:
    toc: true
    toc_depth: 2
    df_print: tibble
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(parallel)
library(readxl)
library(httr)
library(DBI)
library(RSQLite)
library(jsonlite)
```

# Querying APIs in R

## Introduction

The purpose of this vignette is to give a brief overview of accessing APIs within R, writing custom functions that make the process easier, and performing some basic exploratory data analysis (EDA) on the data we use.

The API we are going to utilize is [polygon.io](https://polygon.io/docs/getting-started), which contains some financial data of various stocks, options, and currencies.


### Required Packages

The following packages (and their purpose) will be required:

- `tidyverse`:
- `httr`:


### Basic API Info

Endpoints, scope, etc.


## Accessing the API

What function(s) will do


### Writing Function(s)

Format of input, format of output, etc.

```{r functions}

```


### Function Demonstration

Showing the function works on a few different datasets/endpoints and the format of the returned data

```{r examples}

```


### Analysis Data

Using the function(s) to pull the data for analysis and describing variables

```{r final data}

```


## Exploratory Data Analysis

General overview of what our "response" and/or "predictors" are

### Univariate Summaries

graphs/summaries/etc.

```{r}

```

interesting because...

```{r}

```

insightful because...


### Bivariate Summaries

contingency table(s), grouped summaries, colored/grouped graphs

```{r}

```

```{r}

```


### Trivariate Summaries

sicko mode graphs / summaries

```{r}

```

```{r}

```


### "Conclusions"

Overall / general findings
